
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE, cache = FALSE)
```

```{r include=FALSE}
Sys.setlocale(locale="UTF-8") # Just in case it is not set by the user.
```

```{r }
library(leaflet) 
library(leaflet.extras)
library(leaflet.minicharts)
library(dplyr)
library(DT)

## Load data for all chunks---- 

argenis  = read.csv("data/argenis_places.csv")
etiopicas = read.csv("data/etiopicas_places.csv")
peregrino = read.csv("data/peregrino_places.csv")
persiles = read.csv("data/persiles_places.csv")
semprilis = read.csv("data/semprilis_places.csv")
corpus = read.csv("data/corpus_bizantinas.csv")
piechart_5novels = read.csv("data/piechart_5novels.csv")
piechart_5novels_percentage = read.csv("data/piechart_5novels_percentage.csv")
polonia = read.csv("data/polonia-1618-polygons.csv")
genorodano_no_geo = read.csv("data/Genorodano_itinerary_no-geo.csv")
semprilis_no_geo = read.csv("data/Semprilis_itinerary_no-geo.csv")

```

<br/>

# Introduction

"Digital mapping of fictional places in Spanish Early Modern Byzantine novels" is a project, which adresses the possibilities of digital technologies for quantification and visualization of the narrative space in fictional texts.

Data and scripts are available in the github repository [```editio/mapping.literature```](https://github.com/editio/mapping.literature) and the open-access repository Zenodo [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.1919007.svg)](https://doi.org/10.5281/zenodo.1919007)

# Corpus of novels

Sixteenth and Seventeenth century novels, such as, Lope de Vega’s _Peregrino_ (1604), Miguel de Cervantes’ _Persiles_ (1617), and Enríquez de Zúñiga's _Semprilis_ (1629) belong to the Byzantine genre (so called in the Spanish Studies tradition). Their narrative structure and chronotope is defined by sea travels, pirates, exoticism, shipwrecks, transcultural encounters. The Hellenistic romance, in particular Heliodorus’ _Ethiopica_, had a considerable influence on the genre, as well as the political novel _Argenis_ of John Barclays. They are not Spanish novels, but they are analyzed together for comparison. 

``` {r }

datatable(corpus, style = "bootstrap", options = list(dom = 't'))

```

# Geolocating places

Places have been geolocated using digital gazetteers. An R package [```editio/georeference```](https://github.com/editio/georeference) has been developed in order to automate the process using the Pelagios API. The final data has been refined manually.

```{r }
datatable(semprilis[1:6,c(5,3,4,6,7)], style = "bootstrap", options = list(dom = 't'))
```

# Maps visualizations

To toggle information layers use the menu at the top right corner of the maps. Clicking the code button above the maps displays the scripts in R used to create the visualization.

## Clusters of places

The five novels (each color corresponds to one novel) are clustered as we may expect in the Mediterranean area. Hovering over the clusters, the coverage of the cluster's bounds appears. In Cervantes’ _Persiles_ it shows the characteristic North-South axis; in Heliodorus _Aethiopica_, clearly the Eastern Mediterranean basin, and in Zuñiga’s _Semprilis_, almost the whole Mediterranean area.

``` {r echo = TRUE, out.width = '100%'}

leaflet() %>% 
  
  setView(9, 43, zoom = 4) %>%
  addFullscreenControl(pseudoFullscreen = T) %>%
  addTiles(urlTemplate = "http://server.arcgisonline.com/ArcGIS/rest/services/World_Shaded_Relief/MapServer/tile/{z}/{y}/{x}",
         attribution = 'Base Map &copy; Esri | Project by J.L. Losada Palenzuela',
         group="Base (no labels)",
         option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%

  addProviderTiles ("CartoDB.Positron", group = "Modern") %>% 
  addTiles(urlTemplate = "http://pelagios.org/tilesets/imperium/{z}/{x}/{y}.png",
           attribution = 'Barrington Roman Empire: (CC BY-SA) <a href="http://dare.ht.lu.se">DARE</a> & <a href="http://commons.pelagios.org">Pelagios</a> | Project by J.L.Losada',
           group="Roman Empire",
           option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%  

# Places Heliodoro----  
  
    addCircleMarkers(
    weight = 1,
    data = etiopicas, lng= ~lon, lat= ~lat,
    label = ~searched_name,
    color = "grey",
    stroke = FALSE, fillOpacity = 0.5,
    group = "clusters",
    clusterOptions = markerClusterOptions
    (iconCreateFunction = JS("function(cluster) {
                              return new L.DivIcon({
                              html: '<div style=\"background-color:grey\"><span>' + cluster.getChildCount() + '</div><span>', className: 'marker-cluster'});}")
                              , freezeAtZoom=2
      , polygonOptions = list(color = "grey", opacity = 0.5, fillOpacity = 0.3)
                                )
                  ) %>% 
  
#Places Persiles----  
 
 addCircleMarkers(
    weight = 1,
    data = persiles, lng= ~lon, lat= ~lat,
    label = ~searched_name,
    color = "tomato",
    stroke = FALSE, fillOpacity = 0.5,
    group = "clusters",
    clusterOptions = markerClusterOptions(iconCreateFunction =
                                            JS("
                                               function(cluster) {
                                               return new L.DivIcon({html: '<div style=\"background-color:tomato\"><span>' + cluster.getChildCount() + '</div><span>', className: 'marker-cluster'});}")
                          , freezeAtZoom=2
   , polygonOptions = list(color = "tomato", opacity = 0.5, fillOpacity = 0.3)                          
                                        )
                ) %>% 

# Places Semprilis----
    
    addCircleMarkers(
    weight = 1,
    data = semprilis, lng= ~lon, lat= ~lat,
    label = ~searched_name,
    color = "green",
    stroke = FALSE, fillOpacity = 0.5,
    group = "clusters",
    clusterOptions = markerClusterOptions(iconCreateFunction =
                                            JS("
                                               function(cluster) {
                                               return new L.DivIcon({
                                               html: '<div style=\"background-color:green\"><span>' + cluster.getChildCount() + '</div><span>',
                                               className: 'marker-cluster'
                                               });
                                               }")
                      , freezeAtZoom=2
          , polygonOptions = list(color = "green", opacity = 0.5, fillOpacity = 0.3)
 )  
      ) %>%   
    
# Places Argenis----
  
  addCircleMarkers(
    weight = 1,
    data = argenis, lng= ~lon, lat= ~lat,
    label = ~searched_name,
    color = "white", # aqui para los indiviudales. luego en clusteroptions
    stroke = FALSE, fillOpacity = 0.5,
    group = "clusters",
    clusterOptions = markerClusterOptions(iconCreateFunction =
                                            JS("
                                               function(cluster) {
                                               return new L.DivIcon({
                                               html: '<div style=\"background-color:white\"><span>' + cluster.getChildCount() + '</div><span>',
                                               className: 'marker-cluster'
                                               });
                                               }")
                            , freezeAtZoom=2
      , polygonOptions = list(color = "white", opacity = 0.5, fillOpacity = 0.3)                            
                                          )
                                            ) %>% 
  
# Places Lope----
    
  addCircleMarkers(
    weight = 1,
    data = peregrino, lng= ~lon, lat= ~lat,
    label = ~searched_name,
    color = "orange", # el color aqu'o no importa, se hace en clusteroptions
    stroke = FALSE, fillOpacity = 0.5,
    group = "clusters",
    clusterOptions = markerClusterOptions(iconCreateFunction =
                                            JS("
                                               function(cluster) {
                                               return new L.DivIcon({
                                               html: '<div style=\"background-color:orange\"><span>' + cluster.getChildCount() + '</div><span>',
                                               className: 'marker-cluster'
                                               });
                                               }")
, freezeAtZoom=2
, polygonOptions = list(color = "orange", opacity = 0.5
                        , fillOpacity = 0.3
                        )
                                          )
                                            ) %>%
# Controls

addLayersControl(
  baseGroups = c("Base (no labels)", "Modern", "Roman Empire"),
  overlayGroups = c("clusters"),
  options = layersControlOptions(collapsed = T)) %>%
  
  addLegend(position = "bottomleft", opacity = 0.4, 
          colors = c("orange", "tomato", "green", "white", "gray"), 
          labels = c("«Peregrino» (1604)", 
                     "«Persiles» (1617)", 
                     "«Semprilis» (1629)",
                     "«Argenis» (1621)",
                     "«Eti&oacute;picas» (s. III)"
                     ))
```

<br/>
  
## Places by frequency

The visualisation of places most frequently mentioned offers another spatial perspective: Spain and Poland are the most frequent places in _Semprilis_; Rome, in _Persiles_, as the final destination of the lovers pilgrimage; the Iberian peninsula (Barcelona, Valencia, Toledo) in  _Peregrino_, novel which indeed takes a turn towards a nationalization of the genre.

``` {r echo = TRUE, out.width = '100%'}

# Assign colors by frequency

pal.semp = colorNumeric(
  palette = c("#00cc00", "#004d00"),  
  domain = semprilis$freq)

pal.pere = colorNumeric(
  palette = c("#ffaf1a", "#cc8500"),
  domain = peregrino$freq)

pal.per = colorNumeric(
  palette = c("#ff7d66", "#e62200"),
  domain = persiles$freq)

pal.arg = colorNumeric(
  palette = c("#ffffff", "#cccccc"),
  domain = argenis$freq)

pal.etio = colorNumeric(
  palette = c("#808080", "#1a1a1a"),
  domain = etiopicas$freq)

# Run map

leaflet() %>% 
  
  setView(9, 43, zoom = 4) %>%
  addFullscreenControl(pseudoFullscreen = T) %>%
  addTiles(urlTemplate = "http://server.arcgisonline.com/ArcGIS/rest/services/World_Shaded_Relief/MapServer/tile/{z}/{y}/{x}",
         attribution = 'Base Map &copy; Esri | Project by J.L. Losada Palenzuela',
         group="Base (no labels)",
         option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%
  addProviderTiles ("CartoDB.Positron", group = "Modern") %>% 
  addTiles(urlTemplate = "http://pelagios.org/tilesets/imperium/{z}/{x}/{y}.png",
           attribution = 'Barrington Roman Empire: (CC BY-SA) <a href="http://dare.ht.lu.se">DARE</a> & <a href="http://commons.pelagios.org">Pelagios</a> | Project by J.L.Losada',
           group="Roman Empire",
           option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%

# CircleMarkes for frequencies----
  
addCircleMarkers(data=semprilis, lng = ~lon, lat = ~lat, weight = 1, fillOpacity=0.5,
                   radius = ~sqrt(freq)*3, color = ~pal.semp(freq), label=~searched_name, group="(frq) «Semprilis»")   %>%
  
addCircleMarkers(data=peregrino, lng = ~lon, lat = ~lat, weight = 1, fillOpacity=0.5,
                   radius = ~sqrt(freq)*3, color = ~pal.pere(freq), label=~searched_name, group="(frq) «Peregrino»")   %>%  

addCircleMarkers(data=persiles, lng = ~lon, lat = ~lat, weight = 1, fillOpacity=0.5,
                   radius = ~sqrt(freq)*3, color = ~pal.per(freq), label=~searched_name, group="(frq) «Persiles»")   %>%  
  
addCircleMarkers(data=argenis, lng = ~lon, lat = ~lat, weight = 1, fillOpacity=0.5,
                   radius = ~sqrt(freq)*3, color = ~pal.arg(freq), label=~searched_name, group="(frq) «Argenis»")   %>%  
  
addCircleMarkers(data=etiopicas, lng = ~lon, lat = ~lat, weight = 1, fillOpacity=0.5,
                   radius = ~sqrt(freq)*3, color = ~pal.etio(freq), label=~searched_name, group="(frq) «Eti&oacute;picas»")   %>% 

# Controls  
  
  addLayersControl(
  baseGroups = c("Base (no labels)", "Modern", "Roman Empire"),
  overlayGroups = c("(frq) «Eti&oacute;picas»","(frq) «Argenis»", "(frq) «Persiles»", "(frq) «Peregrino»", "(frq) «Semprilis»"),
  options = layersControlOptions(collapsed = T)) %>%
    
  addLegend(position = "bottomleft", opacity = 0.4, 
          colors = c("orange", "tomato", "green", "white", "gray"), 
          labels = c("«Peregrino» (1604)", 
                     "«Persiles» (1617)", 
                     "«Semprilis» (1629)",
                     "«Argenis» (1621)",
                     "«Eti&oacute;picas» (s. III)"
                     ))
```

## Places clustered by proximity

The heat map represents the geographical proximity of the places mencioned in the novels.

``` {r echo = TRUE, out.width = '100%'}

leaflet() %>% 
  
  setView(9, 43, zoom = 4) %>%
  addFullscreenControl(pseudoFullscreen = T) %>%
  addTiles(urlTemplate = "http://server.arcgisonline.com/ArcGIS/rest/services/World_Shaded_Relief/MapServer/tile/{z}/{y}/{x}",
         attribution = 'Base Map &copy; Esri | Project by J.L. Losada Palenzuela',
         group="Base (no labels)",
         option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%

  addProviderTiles ("CartoDB.Positron", group = "Modern") %>% 
  addTiles(urlTemplate = "http://pelagios.org/tilesets/imperium/{z}/{x}/{y}.png",
           attribution = 'Barrington Roman Empire: (CC BY-SA) <a href="http://dare.ht.lu.se">DARE</a> & <a href="http://commons.pelagios.org">Pelagios</a> | Project by J.L.Losada',
           group="Roman Empire",
           option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%  


addHeatmap(data=peregrino, lng = ~lon, lat = ~lat, blur = 20, max = 0.05, radius = 15, intensity = ~freq, group="(density) «Persiles»") %>%
  
  #gradient = 'Reds'
  
#Persiles----

addHeatmap(data=persiles, lng = ~lon, lat = ~lat, blur = 20, max = 0.05, radius = 15, intensity = ~freq, group="(density) «Peregrino»") %>%
  
#Semprilis----

addHeatmap(data=semprilis, lng = ~lon, lat = ~lat, blur = 20, max = 0.05, radius = 15, intensity = ~freq, group="(density) «Semprilis»") %>%
  
  addLayersControl(
  baseGroups = c("Base (no labels)", "Modern", "Roman Empire"),
  overlayGroups = c("(density) «Persiles»","(density) «Peregrino»", "(density) «Semprilis»"),
  options = layersControlOptions(collapsed = T)) %>%

hideGroup(c("(density) «Persiles»", "(density) «Peregrino»"))
  
```

## Intertextuality. Shared places between novels.

We can have an intertextual glimpse of the common places shared across the genre and its weighted importance in terms of number of mentions for each place, but proportional to the total value in each novel. For example, Rome is mentioned in three novels, but it is more significant in _Persiles_ with a 17.9% of all mentions (click on the piecharts to see the data).  

``` {r echo = TRUE, out.width = '100%'}
colors_share_places <- c("grey", "white", "tomato", "orange", "green")

basemap <- leaflet() %>%
  setView(3, 43, zoom = 4) %>%
  addFullscreenControl(pseudoFullscreen = T) %>%
  addTiles(urlTemplate = "http://server.arcgisonline.com/ArcGIS/rest/services/World_Shaded_Relief/MapServer/tile/{z}/{y}/{x}",
         attribution = 'Base Map &copy; Esri | Project by J.L. Losada Palenzuela',
         group="Base (no labels)",
         option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%

  addProviderTiles ("CartoDB.Positron", group = "Modern") %>% 
  
  addTiles(urlTemplate = "http://pelagios.org/tilesets/imperium/{z}/{x}/{y}.png",
           attribution = 'Barrington Roman Empire: (CC BY-SA) <a href="http://dare.ht.lu.se">DARE</a> & <a href="http://commons.pelagios.org">Pelagios</a> | Project by J.L.Losada',
           group="Roman Empire",
           option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) 

# Data for the popup table

popup.table = paste0("<table>
<div style=\"text-align: center;\">","<font size=2 color=#045FB4>","<a href=",piechart_5novels$geonameid,">", piechart_5novels$lugar, "</a>","</font>","</div> <hr>
           <tr> <td style=\"text-align:right\"><font size=2 color=#404040>Aethiopica:</font></td>
           <td style=\"padding-left: 10px; text-align:right\">", piechart_5novels$Aethiopica," (", round((100 * piechart_5novels_percentage$Aethiopica), digits = 2),"%)","</td> </tr>
            <tr> <td style=\"text-align:right;\"><font size=2 color=black>Argenis:</font></td>
           <td style=\"padding-left: 10px; text-align:right\">", piechart_5novels$Argenis," (", round((100 * piechart_5novels_percentage$Argenis), digits = 2),"%)","</td> </tr>
           <tr> <td style=\"text-align:right\"><font size=2 color=tomato>Persiles:</font></td>
           <td style=\"padding-left: 10px; text-align:right\">", piechart_5novels$Persiles," (", round((100 * piechart_5novels_percentage$Persiles), digits = 2),"%)","</td> </tr>
          <tr> <td style=\"text-align:right\"><font size=2 color=orange>Peregrino:</font></td>
           <td style=\"padding-left: 10px; text-align:right\">", piechart_5novels$Peregrino," (", round((100 * piechart_5novels_percentage$Peregrino), digits = 2),"%)","</td> </tr>
          <tr> <td style=\"text-align:right\"><font size=2 color=green>Semprilis:</font></td>
           <td style=\"padding-left: 10px; text-align:right\">", piechart_5novels$Semprilis," (", round((100 * piechart_5novels_percentage$Semprilis), digits = 2),"%)","</td> </tr>
           </table>")

basemap %>%
  addMinicharts(
    piechart_5novels$lon, piechart_5novels$lat,
    type = "pie",
    chartdata = piechart_5novels_percentage[,c("Aethiopica", "Argenis", "Persiles", "Peregrino", "Semprilis")],
    popup = popupArgs(html = popup.table),
    colorPalette = colors_share_places,
    width = 60 * sqrt(piechart_5novels$total) / sqrt(max(piechart_5novels$total)), legend = F
  )  %>%

 addLayersControl(
  baseGroups = c("Base (no labels)", "Modern", "Roman Empire"),
  options = layersControlOptions(collapsed = T)) %>%
  addLegend(position = "bottomleft", opacity = 0.4, 
          colors = c("orange", "tomato", "green", "white", "gray"), 
          labels = c("«Peregrino» (1604)", 
                     "«Persiles» (1617)", 
                     "«Semprilis» (1629)",
                     "«Argenis» (1621)",
                     "«Eti&oacute;picas» (s. III)"
                     ))
```

## Historical maps and itineraries

The influence of early modern historical cartography is significant to the narrative strategies in these novels. In _Semprilis_, the places of Sevilla, Larache (the old city of Lixa), the river Lucus on the Atlantic coast of Morocco, all share a common fictional space with the Congo because this spatial narrative has been created following an historical map, which represents the Congo close to them, in the inset.

Zoom-out to see the entire itinerary. Pulse circles refer to places without a reference in the geospace, i.e, unknown islands.

On the other hand plotting the historical borders of the Kingdom of Poland (toggle layers in top right corner) and comparing it with the distribution of other places of action (Constantinople, Black Sea, Tartary), confirms the narrative suggestion of an access to the coasts of Poland through the Mediterranean.

``` {r echo = TRUE, out.width = '100%'}

leaflet() %>%
  setView(-8, 35, zoom = 6) %>%
  addFullscreenControl(pseudoFullscreen = T) %>%
  addTiles(urlTemplate = "http://server.arcgisonline.com/ArcGIS/rest/services/World_Shaded_Relief/MapServer/tile/{z}/{y}/{x}", attribution = 'Base Map &copy; Esri | Project by J.L. Losada Palenzuela', group="Base (no labels)", option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%
  addProviderTiles ("CartoDB.Positron", group = "Modern") %>% 
  addTiles(urlTemplate = "http://pelagios.org/tilesets/imperium/{z}/{x}/{y}.png", attribution = 'Barrington Roman Empire: (CC BY-SA) <a href="http://dare.ht.lu.se">DARE</a> & <a href="http://commons.pelagios.org">Pelagios</a> | Project by J.L.Losada', group="Roman Empire", option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11"))  %>%
  
  #  Ortelius Tiles----
  
  addTiles(urlTemplate = "https://editio.github.io/tiles/ortelius/{z}/{x}/{y}.png", attribution = '| Ortelius (Fessae, et Marocchi) &copy; <a href="http://www.bne.es">BNE</a>', group="Ortelius (Fessae, et Marocchi)", option=list(continuousWorld=TRUE,tileSize="256", minZoom="0", maxZoom="10"))  %>%

  # Poland historical borders overlay
  
  addPolygons(data = polonia, lat= ~y, lng= ~x, color = "orange", weight = 1,smoothFactor = 0.5, opacity = 1.0, fillOpacity = 0.5, label = "Rzeczpospolita Obojga Narodów (Poland circa 1618)", group = "Poland circa 1618") %>%

  # Character's routes
  
  addPolylines(data=semprilis_no_geo, lng= ~long, lat= ~lat,  color = "red", group = "Semprilis' route (no geospace)", weight = 4, dashArray = '2,5') %>%  

  addPolylines(data=genorodano_no_geo, lng= ~long, lat= ~lat,  color = "blue", group = "Genorodano's route (no geospace)", weight = 3, dashArray = '3,5') %>%

  # Semprilis y Genorodano. No geospace---- 
  
  # Semprilis
    
    addCircleMarkers(
    color = "red",
    weight = 3,
    data = semprilis_no_geo, lng= ~long, lat= ~lat, 
    popup=paste(sep="", "<b>", semprilis_no_geo$lugar,"</b>", "<br/>","<font size=2 color=#045FB4>","Semprilis' Stop nr: ","</font>", semprilis_no_geo$stops_female),
    radius = 3.5, 
    label = (~lugar),
    stroke = T,
    group = "Semprilis' route (no geospace)", labelOptions = labelOptions(noHide = F, direction = "left", style = list("color" = "red",
      "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
      "font-size" = "10px",
      "border-color" = "rgba(0,0,0,0.5)"
    )))  %>%
  
  # Genorodano
  
  addCircleMarkers(
    color = "blue",
    weight = 1,
    data = genorodano_no_geo, lng= ~long, lat= ~lat, 
    popup=paste(sep="", "<b>", genorodano_no_geo$lugar,"</b>", "<br/>","<font size=2 color=#045FB4>","Genorodano's stop nr: ","</font>", genorodano_no_geo$stops_male),
    radius = 6, 
    label= (~lugar),
    stroke = TRUE,
    group = "Genorodano's route (no geospace)", labelOptions = labelOptions(noHide = F, direction = "right",style = list("color" = "blue",
      "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
      "font-size" = "10px",
      "border-color" = "rgba(0,0,0,0.5)"
    )))  %>%
  
  # Pulse Markers----
  
  addPulseMarkers(data=genorodano_no_geo[c((3),(6:7), (8:9)),], lng= ~long, lat= ~lat, group = "Genorodano's route (no geospace)", icon = makePulseIcon(heartbeat = 2.0, iconSize = 4, color = "blue")) %>%

  addPulseMarkers(data=semprilis_no_geo[c((3),(7),(8:9)),], lng= ~long, lat= ~lat, group = "Semprilis' route (no geospace)", icon = makePulseIcon(heartbeat = 2.0, iconSize = 4, color = "red")) %>%
  
  # Controls  
  
  addLayersControl(
  baseGroups = c("Base (no labels)", "Modern", "Roman Empire" , "No map"),
  overlayGroups = c("Ortelius (Fessae, et Marocchi)", "Poland circa 1618", "Semprilis' route (no geospace)", "Genorodano's route (no geospace)"),
  options = layersControlOptions(collapsed = T)) %>%
  
  hideGroup(c("Poland circa 1618", "Semprilis' route (no geospace)")) %>%
    
  addLegend(position = "bottomleft", opacity = 0.4, 
          colors = c("red", "blue"), 
          labels = c("Semprilis' route", 
                     "Genorodano's route"
                     ))
```

## Descriptive decoration in historical maps

Maps present decorations such as monsters, ships, battles, etc., which convey additional semantic layers to maps. This map by the famous cartographer Abraham Ortelius represents a battle with sea raiders operating in the Gulf of Cádiz at the southwestern coast of the Iberian Peninsula. Cervantes thematizes its use in the _Persiles_ when the false captives relate their (false) history leaning deictically on a canvas where a ship (a galiot) is drawn.

> "se atreven a pasar el plus ultra de las colunas de Hércules, y a acometer y robar las apartadas islas, que, por estar rodeadas del inmenso mar Océano, pensaban estar seguras, a lo menos de los bajeles turquescos. Este bajel que aquí veis, reducido a pequeño, porque lo pide así la pintura, es una galeota de ventidós bancos" (Cervantes, _Los trabajos de Persiles y Sigismunda_, Carlos Romero (ed.), Madrid, 2004: III, 10, 529)  

``` {r echo = TRUE, out.width = '100%'}

leaflet() %>%
  setView(-6.5, 36.6, zoom = 10) %>%
  addFullscreenControl(pseudoFullscreen = T) %>%
  addTiles(urlTemplate = "http://server.arcgisonline.com/ArcGIS/rest/services/World_Shaded_Relief/MapServer/tile/{z}/{y}/{x}", attribution = 'Base Map &copy; Esri | Project by J.L. Losada Palenzuela', group="Base (no labels)", option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11")) %>%
  addProviderTiles ("CartoDB.Positron", group = "Modern") %>% 
  addTiles(urlTemplate = "http://pelagios.org/tilesets/imperium/{z}/{x}/{y}.png", attribution = 'Barrington Roman Empire: (CC BY-SA) <a href="http://dare.ht.lu.se">DARE</a> & <a href="http://commons.pelagios.org">Pelagios</a> | Project by J.L.Losada', group="Roman Empire", option=list(continuousWorld=TRUE, tileSize="256", minZoom="0", maxZoom="11"))  %>%
  
  #  Bahía de Cádiz----
addTiles(urlTemplate = "https://editio.github.io/tiles/ortelius_cadiz/{z}/{x}/{y}.png",
         attribution = '| Mapa de Ortelius: @ <a href="http://www.bne.es">BNE</a>', 
         group = "Bahía de Cádiz",
         option=list(continuousWorld=TRUE, tileSize="256", minZoom="3", maxZoom="13"))  %>%

# Controls  
  
  addLayersControl(
  baseGroups = c("Base (no labels)", "Modern", "Roman Empire" , "No map"),
  overlayGroups = c("Bahía de Cádiz"),
  options = layersControlOptions(collapsed = T))

```

# Papers and dissemination

Parts of project were presented at different conferences: 

- _Mapping the Text. Humanistic Scholarship on Spatiotemporalities within Cultural & Æsthetic Objects_, 20-22.04.2018, New York University. See the slides ["Digital mapping of fictional places in Spanish Early Modern Byzantine novels"](http://editio.github.io/slides/literary.maps) ).

- _Computational Methods in the Humanities_, 4-5.06.2018, University of Lausanne.

- _Cervantes, su obra y su tiempo. Cuarto centenario (1616-2016)_, 19-21.12.2016, University of Navarra. See the slides ["En la estela del _Persiles_. Cartografía y verosimilitud en Historias de las fortunas de _Semprilis y Genorodano_"](http://editio.github.io/slides/semprilis-mapas) ).

Related papers will be published in _Artnodes. Journal on art, science and technology_ (forthcoming 2019) and _Revista de Literatura_ (forthcoming 2020). 

# How to cite
 
 Losada Palenzuela, José Luis (2018), _Digital mapping of literature_, [<http://editio.github.io/mapping.literature>](http://editio.github.io/mapping.literature)
 
<br/>
