---
title: ""
output: 
  html_document:
    toc: true
    code_folding: NULL
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE, cache = FALSE)
```

```{r,echo=FALSE}
load(file="data/maps.RData")
```

```{r}
#knitr::read_chunk('scripts/spatial_net_functions.R')
source('scripts/spatial_net_functions.R', echo = FALSE)
source('scripts/spatial_net.R', echo = F)
source('scripts/spatial_net_map.R', echo = F)
source('scripts/spatial_net_viz.R', echo = F)
source('scripts/leafsync.R', echo = F)
```

<br/>

A spatial feature of the genre models —Heliodorus’ _Ethiopica_, Cervantes' _Persiles_— is the presence of a geographical division between **core and periphery**. This is an atempt to automatically compute the geographical periphery and center of the places mentioned in a text using network metrics (eigenvector values). 

Some places have been refined again manually, e.g., locations like Thule has been added in _Persiles_.

# Voronoi diagrams

```{r}
plot2_voro

plot4_map
```

# Delaunay triangulation

```{r}
plot1_both
plot3_dela
```

# Graph from a TIN

```{r fig1, message=F}
par(mar = c(0,0,0,0))

plot(grafo,
      vertex.size= ifelse(V(grafo)$ec == 1, 8, 3),
      vertex.label = NA,
      edge.width=E(grafo)$weight/3,
      vertex.frame.color = "red",
      vertex.color = ifelse(V(grafo)$ec == 1, "blue", "red")
      )
```

# Metrics (eigenvector centrality)

```{r}
datatable
```

# Metrics plot on a map
```{r}
m1
```

# Comparation of novels
```{r}
retablo
```

<br/>
